% ============================================================
% Appendix
% ============================================================



\section{Full Proofs}
\label{app:proofs}

This appendix collects the technical material that supports the main
paper's central claims but would be too detailed for the main
narrative. We begin with proof sketches for the sampler properties,
then report the composability analysis, hyperparameter sensitivity, and
per-conversation breakdowns referenced in the main text.

\subsection{Proof of Lemma~\ref{lem:dist-preserve}}
The key idea is the standard binning argument underlying
distribution-preserving keyed selection. Integer-quantize
$\probdist$ into $N$ buckets, apply a cyclic shift determined by a
uniform PRF output, and return the candidate whose interval contains
the shifted point. Marginalizing over the uniform shift recovers the
original interval mass for every candidate, so the keyed pick has the
same marginal distribution as $\probdist$.

\subsection{Proof of Lemma~\ref{lem:cascade}}
Distinct internal decisions use distinct context strings and therefore
distinct nonces $\nonceT = \mathrm{PRF}(K, \ctx)$. Under the standard
PRF assumption, these nonces are computationally indistinguishable from
independent random draws to an adversary without the key. The keyed
picks across a cascade of LLM calls therefore behave independently for
verification purposes, and the expected embedding capacity adds across
valid decision points.

\subsection{Proof of Lemma~\ref{lem:backend-invariant}}
This is an immediate corollary of
Lemma~\ref{lem:dist-preserve}. The marginal expression depends only on
the candidate distribution $\probdist$ and the secret key $K$.
Backend-specific implementation details affect how the candidate set is
enumerated, but once $(\candset, \probdist, \ctx)$ is fixed, the
sampling guarantee is backend-agnostic.

\section{Metrics}
\label{app:metrics}

\input{latex/tables/metric_definitions}
Table~\ref{tab:metric-defs} consolidates the definitions of all metrics
used in this paper, grouped by the research question they address: utility
(RQ1), watermark reliability (RQ2), capacity (RQ3), tamper detection (RQ4),
and memory integrity (RQ5). For each metric we list its reporting unit and,
where relevant, the carrier-level decomposition along
$\carrier \in \{\text{update\_target},\ \text{link\_target},\
\text{semantic\_realization}\}$. We refer the reader to this table
throughout Sections~\ref{sec:rq1}--\ref{sec:rq5} for precise definitions.


\section{Composability Experiment}
\label{app:composability}
We evaluate three configurations --- memory-only, content-only, and
both --- under the R1 protocol of \S\ref{sec:rq3}. The purpose is to
check whether memory-level attribution and output-level watermarking
remain operational when deployed simultaneously. We report
(i) memory decode rate, (ii) content watermark detection, and
(iii) joint utility. Table~\ref{tab:composability} summarizes the
comparison.

\input{tables/composability}

\section{Hyperparameter Sensitivity}
\label{app:hyperparam}
We sweep $K \in \{2, 4, 8\}$ and
$T_{\text{enum}} \in \{0.5, 0.7, 1.0\}$ to quantify the trade-off
between candidate diversity and embedding stability. We report realized
bits per decision together with the corresponding utility deltas, so
that the reader can see how much attribution capacity is obtained per
unit of behavioral perturbation.

\input{tables/hyperparam_sweep}

\section{Per-Conversation Breakdown}
\label{app:per-conv}
Finally, we provide the per-conversation breakdown for all
10 \locomo{} conversations. This table is included only in the
appendix because its role is diagnostic rather than conceptual: it
shows that the aggregate trends in the main text are not driven by a
single conversation outlier.

\input{tables/per_conversation}
