% ============================================================
% Experiments — three-pillar story.
%
% Order: RQ1 → RQ2 → RQ3 → RQ4 → RQ5
%   §4.1 Setup
%   §4.2 RQ1 — Utility Preservation
%   §4.3 RQ2 — Capacity
%   §4.4 RQ3 — In-Record / Snapshot-Only Verification (HEADLINE)
%   §4.5 RQ4 — Robustness against 9 memory-lifecycle attacks
%   §4.6 RQ5 — Memory Integrity
%
% Rationale: standard watermark-paper layout. RQ1 is the gate
% ("watermark doesn't break memory"); only once that gate is
% passed does RQ3's attribution claim become meaningful. RQ3 is
% still flagged as the headline experiment in its subsection
% title and in the contributions list.
% ============================================================
\section{Experiments}
\label{sec:exp}
Our experiments evaluate whether \memmark{} enables robust attribution without degrading the memory system. We investigate if it preserves downstream utility (\textbf{RQ1}), contains sufficient entropy to embed payloads (\textbf{RQ2}), survives snapshot-only regimes like memory leaks (\textbf{RQ3}), resists nine distinct memory-lifecycle attacks (\textbf{RQ4}), and avoids latent damage to the memory lifecycle itself (\textbf{RQ5}). 

% ============================================================
\subsection{Setup}
\label{sec:setup}
% ============================================================
\paragraph{Backends \& Models.}
Our evaluation spans two memory backends: \amem{} \cite{amem}, an agentic note-centric memory system with structured memory notes, dynamic indexing, inter-memory linking, and memory evolution, and \graphiti{} \cite{graphiti}, a temporal graph-based memory system that tracks evolving entities, relations, provenance, and time-aware updates. For model backbones, we use three LLMs: DeepSeek-V4-Pro \cite{deepseekai2026deepseekv4}, 
Qwen3.6-flash \cite{qwen36_35b_a3b} and GLM-5.1 \cite{glm5team2026glm5vibecodingagentic}. Candidate enumeration uses
$T_{\text{enum}}{=}0.7$, candidate scoring uses
$T_{\text{score}}{=}0.0$, and JSON mode is enabled throughout.

\paragraph{Benchmark.}
We evaluate on the \locomo{} benchmark \cite{maharana-etal-2024-evaluating} which is a standard testbed for long-term conversational memory consisting of ten multi-session dialogues with QA and event-summarization annotations across five reasoning categories (single-hop, multi-hop, temporal, commonsense, adversarial).


\paragraph{Baselines.}
We compare against five baselines chosen to isolate different parts of
the claim. \texttt{no-watermark} gives the utility upper bound.
\texttt{random-replace} measures the false-positive and wrong-key
floor by replacing candidates without a secret key.
\texttt{signed-metadata-only} keeps the audit trace and signed anchors
but disables keyed embedding, thereby isolating the marginal value of
the watermark over provenance metadata alone in the R3 setting.
We also adopt \kgmark{} \cite{kgmark} which is the direct structured-memory baseline on \graphiti{}. 

\paragraph{Metrics.}
We evaluate along five axes, aligned with RQ1--RQ5; full definitions are in Appendix~\ref{app:metrics} (Table~\ref{tab:metric-defs}). For \emph{utility preservation} (RQ1) we report \locomo{} F1 and BLEU-1. For \emph{capacity} (RQ2) we report two views: opportunity-side capacity (average candidate set size and per-decision entropy $H(p_t)$) and realized capacity (total embedded payload and its per-carrier composition). For \emph{in-record/snapshot-only verification} (RQ3) we report bit recovery rate, full-message decode success rate, FPR on unwatermarked sessions, and wrong-key acceptance rate across the R1/R2/R3 regime hierarchy. For \emph{robustness against nine memory-lifecycle attacks} (RQ4) we report pre- and post-attack bit recovery, commitment and Merkle-proof verification failure rates, and missing-leaves rate, all per attack type and strength \footnote{Attack descriptions are in Appendix \ref{app:attack}}. For \emph{memory integrity} (RQ5) we report snapshot size, carrier distribution, QA-time evidence-grounded retrieval recall, and write-failure rate. All reliability and capacity metrics are additionally broken down by carrier $\carrier \in \{\text{update\_target}, \text{link\_target}, \text{semantic\_realization}\}$. %todo:attack description in appendix



% ============================================================
\subsection{RQ1 — Utility Preservation}
\label{sec:rq1}
% ============================================================
Utility preservation is the first requirement for any practical memory
watermark. Table~\ref{tab:main-results} therefore compares \memmark{}
with unwatermarked execution and the ablated controls on \locomo{}
across the two backends.

The main pattern is that \memmark{} does not induce uniform utility
collapse and remains competitive with the non-watermarked and ablated
controls. On \amem{}, it improves over \texttt{No-WM} for
\textsc{Qwen3-6-Flash} ($0.2850 \rightarrow 0.3044$ overall F1) and
\textsc{GLM-5} ($0.2958 \rightarrow 0.3181$), while remaining in the
same range as the stronger controls for \textsc{DeepSeek-V4-Pro}. On
\graphiti{}, the picture is more mixed: \memmark{} is not uniformly
best, but it stays within the same overall operating regime as the
comparison methods rather than failing sharply. Overall, these
results suggest that memory-level attribution can be added without
severe degradation in downstream QA quality.

\input{tables/main_results}
% \input{figures/per_category} % todo:move to appendix

% ============================================================
\subsection{RQ2 — Capacity}
\label{sec:rq2}
% ============================================================

After establishing utility preservation, we verify the host channel's watermark capacity, which is upper-bounded by the entropy of the LLM's candidate set ($K{=}4$). Figure~\ref{fig:entropy-hist} demonstrates robust per-decision entropy $H(p_t)$: near-deterministic decisions ($H{<}0.2$) are nonexistent, and mean entropy consistently exceeds one bit across all carriers (update $1.10$, link $1.14$, semantic $1.36$), operating well within the theoretical maximum of $\log_2 4 = 2$. This confirms the candidate distribution retains substantial selection freedom rather than collapsing onto a single option, yielding a viable, non-degenerate channel. We next confirm this nominal capacity translates to fully recoverable bits end-to-end. Figure~\ref{fig:carrier_bits} shows the complete $40$-bit payload is realized across all model--backend combinations on \locomo{}. Carrier distribution is governed by the backend's evolve pipeline rather than the specific LLM: \amem{} distributes bits evenly across carriers, whereas \graphiti{} concentrates the bulk of the payload in semantic realization due to its higher decision density. Ultimately, \memmark{}'s binning sampler efficiently consumes this available entropy to guarantee total payload recovery.


\input{figures/entropy_histogram}
\input{figures/carrier_bits}

% ============================================================
\subsection{RQ3 — Snapshot-Only / Partial-Log Verification}
\label{sec:rq3}
% ============================================================


%%comment: R1 和 R3 的 MEMMARK 行数值完全一样(都 1.00)。 R3 的 in-record sidecar(reveal\_t + 签名 anchor + 每片 leaf 的 inclusion proof)把外部 log 的内容复制进了 snapshot 里,所以正确 key + 无攻击 时 R1 和 R3 在"verifier 拿得到什么"这件事上等价 → 数字必然一样。 这一节的 headline 不是 "R3 比 R1 强",而是 "R3 在没有外部 log 的 情况下达到了 R1 的水平"。R3 这一行单独抽出来做一张柱状图,横轴四个 method (MEMMARK / signed-metadata / wrong-key / Action-layer),纵轴 bit recovery。再画一条水平虚线在 y = 1/K(随机基线),标注上去 —— 这样三个对照 一眼就能看到:watermark 的边际价值(vs 0)、PRF 真在锁(vs 0.15 低于随机)、trajectory-bound 在 snapshot-only 设定下直接失效(N/A)
RQ3 tests the paper's central
claim under the hardest and most practically relevant forensic regime.
When memory is leaked, distilled, or migrated across systems, the
visible action trajectory is typically unavailable. A watermark that
requires the original trajectory is therefore unusable precisely in the
setting where attribution matters most. The defining question for
\memmark{} is whether attribution survives when verification has access
only to the memory state and the cryptographic residue embedded with
it.

We study three progressively weaker evidence regimes.
\textbf{R1} assumes access to the full external audit log and serves as
an upper bound. \textbf{R2} assumes that only a subset of the external
log is retained, with keep ratio
$r \in \{0.1, 0.3, 0.5, 0.7, 0.9\}$. \textbf{R3} removes the external log entirely and requires the verifier to
operate only on the memory snapshot, the in-record reveal sidecar, and
the anchored session header described in
\S\ref{sec:method-audit}. Table~\ref{tab:verification} summarizes the
regime hierarchy, and Figure~\ref{fig:r2-curve} shows the R2 sweep.

\input{tables/verification} 

The main result is that \memmark{} remains fully attributable in the
snapshot-only setting. In R1, it recovers all 40 embedded bits with
commitment pass rate 1.0. In R2, recovery degrades smoothly with the
retained external-log fraction: bit recovery is 0.075, 0.200, 0.600,
0.625, and 0.850 at keep ratios
$r = 0.1, 0.3, 0.5, 0.7, 0.9$, while root matching and
anchor-signature validation remain 1.0 throughout. In R3, where the
verifier receives only the snapshot and the co-located sidecar,
\memmark{} again recovers all 40/40 bits with valid root and signature
checks. The carrier-level R3 recovery is likewise complete: 13/13
\texttt{link\_target} bits, 16/16
\texttt{semantic\_realization} bits, and 11/11
\texttt{update\_target} bits are recovered.

Two controls are critical for interpreting this result. First, the
\texttt{signed-metadata-only} baseline retains valid anchors in every
regime but recovers 0 bits throughout, showing that the attribution
signal does not come from the sidecar metadata by itself. Second, the
wrong-key R3 control recovers only 0.150 of payload bits and fails
anchor-signature validation, whereas the correct-key R3 run recovers
1.000. Together, these controls rule out a trivial verifier that
merely accepts any well-formed reveal record. Successful R3
attribution requires the hidden keyed state-evolution choices to match
the verifier's secret.


\input{figures/r2_curve}

% ============================================================
\subsection{RQ4 — Robustness Against Memory-Specific Attacks}
\label{sec:rq4}
% ============================================================
The R3 result is only useful if it survives the kinds of mutations that
real memory systems undergo over time. We therefore stress-test the
watermark under nine memory-specific attacks spanning both content
rewrites and structural modifications, each applied at strengths
$0.1$, $0.3$, and $0.5$. The goal is not merely to measure post-attack
recovery, but to understand how recovery and tamper evidence trade off
under different failure modes.

Before attack, recovery is 1.0. Content-rewriting attacks show the
expected recovery--detection tradeoff. For example,
\texttt{manual\_edits} yields post-attack recovery of 0.900, 0.750,
and 0.625 as attack strength increases, with corresponding tamper
detection rates of 0.100, 0.300, and 0.499.
\texttt{paraphrase\_rewrite} is more damaging, dropping recovery to
0.825, 0.675, and 0.350. KG-style perturbations are similarly severe
at high strength: both \texttt{edge\_relabel} and
\texttt{subgraph\_reanchor} fall to 0.250 recovery at strength 0.5.
By contrast, purely structural deletion-style operations are less
harmful to the surviving watermark trace. \texttt{pruning} and
\texttt{dedup} preserve 1.000 recovery across all tested strengths,
while \texttt{poisoning} remains comparatively robust at 0.909, 0.816,
and 0.741.

This split is consistent with the audit design in
\S\ref{sec:method-audit}. Structural attacks remove or reorganize
leaves, but they do not necessarily invalidate the surviving
per-record proofs, so recovery can remain smooth when enough valid
trace survives. Content tampering, by contrast, directly perturbs the
committed record bytes and therefore causes recovery to fall in tandem
with commitment-level failures. Figure~\ref{fig:attack-signal-split}
visualizes this separation between content-tamper and structural-tamper
signals. The broader conclusion is that memory-lifecycle mutations do
not create a single monolithic failure mode: some attacks erase signal,
some expose themselves cryptographically, and many do both.

\input{tables/robustness}
\input{figures/attack_signals}

% ============================================================
\subsection{RQ5 — Memory Integrity}
\label{sec:rq5}
% ============================================================
Finally, we ask whether the watermark changes the memory system's
internal behavior in undesirable ways even when end-task utility looks
stable. This is a stricter integrity question: a method could preserve
QA metrics yet still skew which writes are made, increase write-path
instability, or degrade evidence-grounded retrieval.

Table~\ref{tab:integrity} therefore reports only integrity signals that
can be computed end-to-end on \locomo{}. All four \amem{} runs produce
the same 419 records, with no detected duplicates or contradictions in
the JSON snapshots. \memmark{} yields a carrier distribution of
213:200:426 (update:link:semantic), compared with 122:298:421 for
\texttt{no-watermark}. Importantly, this shift reflects where the
keyed sampler found admissible decision points rather than a change in
final snapshot size. Evidence-grounded retrieval recall is 0.585 for
\memmark{}: below \texttt{no-watermark} (0.660), but above
\texttt{signed-metadata-only} (0.552) and
\texttt{random-replace} (0.563). Write failures remain identical to the
unwatermarked run (4), indicating that the watermark layer does not
introduce additional write-path instability.

Taken together with RQ1, these signals suggest that \memmark{} alters
where attribution is embedded without visibly corrupting the memory
lifecycle itself. This distinction is important for long-term memory
settings, where latent write-path damage can accumulate long before it
appears in a single task metric.

\input{tables/integrity}


