% ============================================================
% Introduction — README §1
%   §1.1 Provenance-aware memory is a hot topic, bound to honest writers
%   §1.2 But the attacker is the writer
%   §1.3 Attribution as behavioral trace, not field
%   §1.4 Headline: In-Record Attribution Verification (R3)
%   §1.5 Contributions
% ============================================================
\section{Introduction}
\label{sec:intro}

% ---------- §1.1 honest-writer premise ----------
\paragraph{Provenance-aware memory is bound to honest writers.}
\todo{Open with the active line of work (\tiermem, MemOS MemCube,
enterprise stamping) and the structural assumption they share:
metadata fields are facts because the writer is trusted. Cite
\tiermem{} and MemOS preprints. End with the failure mode: once
memory leaves the original system, that assumption collapses.}

% ---------- §1.2 attacker is the writer ----------
\paragraph{The attacker is the writer.}
\todo{Enumerate what the snapshot-holder can do: stamp ownership,
erase \texttt{source\_agent} / \texttt{tenant\_id}, fabricate
provenance chains, rewrite the memory system's own changelog
(\amem{} evolve history, \graphiti{} fact-invalidation chain) since
they sit in the same store. Conclude: any field-based provenance is
ill-defined in adversarial storage. Token-level LLM watermarks and
agent action-layer watermarks don't fill the gap — they assume the
trajectory survives, which it doesn't in the most forensically
valuable case.}

% ---------- §1.3 attribution = behavioral trace ----------
\paragraph{Attribution as behavioral trace, not field.}
\todo{Frame the three carriers (update / link / semantic) as latent
decisions every memory write makes. Properties: (i) non-trivial
candidate space, (ii) equivalent choices don't break utility, (iii)
the choice itself is readable from the snapshot. Then embed a PRF
key into the sequence of keyed picks: $\nonceT = \text{PRF}(K, \ctx)$
selects which equivalent candidate is realized. Reads as ``\emph{the
writer's key can be replayed against the snapshot}'' rather than
``\emph{we say so}''.}

% ---------- §1.4 R3 headline ----------
\paragraph{Headline: In-Record Attribution Verification (R3).}
\todo{Define the three regimes R1 (full external log) / R2 (partial
log) / R3 (snapshot only). R3 is the headline: bind each keyed
selection into a per-decision commitment, anchor the per-session
Merkle root + signature directly inside the snapshot, and store the
reveal record co-located with the memory record itself. Forward to
\S\ref{sec:rq3} for the numbers.}

% ---------- §1.5 contributions ----------
% NOTE: order reflects the attribution-first reframe.
% Old order led with backend taxonomy; new order leads with R3
% headline so the reader sees the central claim of the paper before
% the engineering pieces that enable it.
\paragraph{Contributions.}
\begin{enumerate}[topsep=2pt,itemsep=2pt,leftmargin=*]
\item \textbf{In-Record Attribution Verification (R3).}
  \todo{One sentence on the three-regime hierarchy
  (\S\ref{sec:rq3}) and the headline finding: \memmark{} recovers
  the full payload from snapshot alone (bit\_recovery $= 1.00$),
  while a \texttt{signed-metadata-only} baseline drops to $0$ and
  the wrong-key control drops to $0.15$. This is the central
  claim — no prior watermark or provenance scheme provides
  snapshot-only attribution under the adversarial-storage threat
  model.}
\item \textbf{Cryptographic audit trace for memory evolution.}
  \todo{One sentence on commitment + per-session Merkle log +
  signed in-snapshot anchor + per-leaf inclusion proof
  (\S\ref{sec:method-audit}). The inclusion proof is what makes R3
  smooth under structural attacks (pruning / dedup / poisoning) —
  each leaf verifies independently against the anchor, so removing
  leaves does not invalidate the survivors.}
\item \textbf{Backend-invariant evolve carrier taxonomy.}
  \todo{One sentence on the $\langle \carrier, \candset, \probdist,
  \ctx \rangle$ formalization (\S\ref{sec:method-formal}) and the
  $\sim$50–100-line native-API wrapper that suffices to instrument
  any memory backend exposing the three carriers.}
\item \textbf{LLM-call-boundary interception with self-reported
  candidate weights.} \todo{One sentence on the \agentmark-style
  1-call enumeration (\S\ref{sec:method-interception}) that
  produces $(\candset, \probdist)$ from an unbounded structured-JSON
  LLM output, enabling distribution-preserving keyed sampling on a
  third-party SDK without touching its internals.}
\item \textbf{Memory-specific evaluation under a controlled
  utility trade-off.} \todo{One sentence on the nine memory-lifecycle
  attack threat model (\S\ref{sec:rq4}) and the framing of
  utility cost as a quantified trade-off rather than a
  preservation claim: \memmark{} pays $\sim$7\% relative F1
  (\S\ref{sec:rq1}) — comparable to KGW and SynthID's 5–15\% — in
  exchange for the attribution guarantees of contribution (i),
  none of which the alternatives provide in the snapshot-only
  setting.}
\end{enumerate}

% ---------- pointer figure ----------
\input{figures/overview}
