
% ============================================================
\section{Problem Formulation}
\label{sec:problem}
% ============================================================

% ---------- 3.1 ----------
\subsection{Preliminaries and Notation}
\label{sec:problem-prelim}
We consider a long-term memory backend that evolves over $T$ turns.
At each turn $t\in\{1,\ldots,T\}$ the backend receives an incoming event
$e_t$ and the current snapshot $M_t$, and produces an updated snapshot
$M_{t+1}$. Crucially, this transition is not monolithic: it is realized
through one or more latent state-evolution choices that are semantically
admissible yet not uniquely determined.

We formalize each such choice as a memory-evolution decision
\begin{equation}
\mathrm{decision}_{t} = \langle \mathrm{carrier}_{t},\ \mathrm{candset}_{t},\ \mathrm{probdist}_{t},\ \mathrm{ctx}_{t} \rangle
\end{equation}
where $\carrier_t$ is the type of state-transition freedom (carrier),
$\candset_{t}=\{c_{t}^{1},\ldots,c_{t}^{k_{t}}\}$ is the candidate set,
$\probdist_t\in\Delta(\candset_t)$ is the backend's preference distribution
over $\candset_t$, and $\ctx_t$ is the decision context that the embedder
and verifier can both reconstruct.

We assume the backend induces a latent preference distribution
$\probdist_t^\star$ over $\candset_t$, and that the baseline backend would
commit a candidate $c_t\sim\probdist_t^\star$. In our setting we elicit an
explicit estimate $\probdist_t$ of $\probdist_t^\star$ from the model's
self-reported weights, and define distribution preservation with respect
to $\probdist_t$. To avoid shifting the backend's writing policy we require
per-decision distribution preservation:
\begin{equation}
\hat c_t \sim \probdist_t \quad \text{for all } t\in\{1,\ldots,T\}.
\end{equation}

\paragraph{Decision--commit separation.}
\memmark{} observes and modifies only $(\candset_t, \probdist_t, \ctx_t)$;
candidate generation and the commit step
$\textsc{apply\_selected}(M_t,\hat c_t)\to M_{t+1}$ remain on the native
backend path. The watermark therefore does not manipulate write operations
themselves, only the selection that feeds into them.

% ---------- 3.2 ----------
\subsection{Memory-Evolution Channel and Distribution-Preserving Coding}
\label{sec:problem-channel}
We model the backend's decision stage as sampling from a time-varying
discrete channel. At each decision point $t$, the channel is characterized
by the latent preference distribution $\probdist_t^\star$ over $\candset_t$.
An unwatermarked backend would commit an implicit choice
$c_t\sim\probdist_t^\star$. To embed a provenance payload, the watermarked
backend replaces native sampling with a distribution-preserving encoder:
\begin{equation}
\hat c_t \leftarrow \mathsf{Enc}(\probdist_t, r_t)\in\candset_t,
\end{equation}
where $r_t$ denotes shared per-decision sampling randomness. The embedder
and verifier share a secret key $K$ so that the verifier can reproduce
$r_t$ from logged context.

\paragraph{Keyed pseudorandomness.}
Per-decision randomness is derived from $K$ and the decision context as
\begin{equation}
\nonceT \leftarrow \mathrm{PRF}(K,\ \ctx_t),
\end{equation}
and a pseudorandom generator seeded by $\nonceT$ produces the sampling
stream consumed by $\mathsf{Enc}$. Because $\ctx_t$ is reconstructible from
the surviving record, the verifier replays the same keyed pick at
verification time.

% ---------- 3.3 ----------
\subsection{Threat Model}
\label{sec:problem-threat}
We consider provenance verification for long-term memory backends, where
the watermark is embedded at write time but verified later, possibly after
the backend has undergone structural modifications and after the external
audit log has been partially or fully lost.

\paragraph{Structural attacks on the backend.}
A long-term memory store is routinely subject to poisoning and
compaction~\cite{chen2024agentpoison,amemguard,memorygraft}. These
operations remove, merge, or rewrite individual records but cannot,
without $K$, forge a commitment that opens consistently against the
anchored Merkle root. We accordingly model the dominant attack effect as
removing a subset of leaves rather than fabricating valid ones. Let
$\mathcal{I}\subseteq\{1,\ldots,T\}$ denote the index set of decisions
whose records survive.

\paragraph{Verification regimes.}
We consider three regimes reflecting realistic deployments:
\begin{itemize}
  \item \textbf{R1 (full external log):} the verifier holds all
  per-decision commitments and the complete Merkle tree.
  \item \textbf{R2 (partial external log):} only a subset $\mathcal{I}$ of
  commitments survives, due to truncation, retention limits, or partial
  loss.
  \item \textbf{R3 (snapshot only):} the external log is unavailable;
  verification relies on the memory snapshot plus an in-record sidecar
  carrying reveal data and the anchored header.
\end{itemize}

% ---------- 3.4 ----------
\subsection{Objectives}
\label{sec:problem-objectives}
We formalize two objectives, with per-decision distribution preservation
$\hat c_t\sim\probdist_t$ enforced as a hard constraint that supports
utility preservation.

\paragraph{Utility preservation.}
Watermarking should not degrade the backend's memory quality on
downstream memory-conditioned tasks. Writing $M_T$ and $\hat M_T$ for the
baseline and watermarked snapshots after $T$ turns, we require
\begin{equation}
\Big|\mathbb{E}[\mathcal{U}(\hat M_T)] - \mathbb{E}[\mathcal{U}(M_T)]\Big|
\le \varepsilon_{\mathcal{U}},
\end{equation}
where $\mathcal{U}(\cdot)$ denotes any downstream utility metric of the
memory backend. This objective motivates restricting carriers to choices
whose candidates are semantically equivalent.

\paragraph{Robust attribution.}
Under the three regimes, the verifier should recover the provenance signal
whenever a sufficient subset of decisions survives. In R1 and R2, given
surviving index set $\mathcal{I}$, we require
\begin{equation}
\Pr\!\left[\mathsf{Verify}\big(\{\hat c_t,\probdist_t,\ctx_t\}_{t\in\mathcal{I}};\ K\big)=1\right]
\ge 1-\delta(|\mathcal{I}|),
\end{equation}
with the failure probability $\delta(|\mathcal{I}|)$ decaying smoothly as
$|\mathcal{I}|$ grows. In R3, the same guarantee is required against the
in-record sidecar instead of the external commitments.


% ============================================================
\section{MemMark}
\label{sec:method}
% ============================================================
\input{figures/pipeline}

\subsection{Overview}
\label{sec:method-overview}
Figure~\ref{fig:pipeline} illustrates \memmark{} in a memory backend.
\memmark{} operates at the decision layer defined in
Section~\ref{sec:problem-prelim} and is organized into three components:
a carrier taxonomy with a backend adapter
(Section~\ref{sec:method-carrier}) that exposes decision points uniformly
across backends; a distribution-preserving sampler
(Section~\ref{sec:method-sampler}) that performs the keyed pick; and a
cryptographic audit trace (Section~\ref{sec:method-audit}) that supports
verification under R1--R3.


% ---------- 4.2 carrier taxonomy + adapter ----------
\subsection{Carrier Taxonomy and Backend Adapter}
\label{sec:method-carrier}
\input{tables/carrier_taxonomy}

Table~\ref{tab:carrier-taxonomy} lists our three carriers. The
update-target carrier changes which existing object is modified; the
link-target carrier changes which prior object is connected to new
evidence; and the semantic-realization carrier changes the surface form
encoding the same fact. Each captures genuine decision freedom in both
backends while satisfying the semantic-equivalence requirement of
Section~\ref{sec:problem-objectives}.

Connecting a backend to \memmark{} requires only a minimal adapter with
three hooks:
\textsc{enumerate\_candidates}$(M_t,e_t,\carrier)\to\candset$,
\textsc{score\_candidates}$(\candset,\ctx)\to\probdist$, and
\textsc{apply\_selected}$(M_t,c^*)\to M_{t+1}$. They expose the choice
space, rank it, and commit the selection. The same abstraction yields the
carrier-level quantities reported throughout the paper: average set size
$|\candset|$, entropy $H(\probdist)$, and acceptance rate (fraction of
points with $|\candset|\ge 2$). Because the adapter is the only
backend-specific code path, a single watermarking layer operates over
structurally distinct systems such as \amem{} and \graphiti{}.


% ---------- 4.3 sampler ----------
\subsection{Distribution-Preserving Behavioral Watermark}
\label{sec:method-sampler}

\paragraph{Boundary interception.}
Each internal LLM call in the backend already constitutes an
$\varepsilon$-equivalent decision point: under standard decoding, the
backend would accept several nearby alternatives without altering the
semantic role of the write. \memmark{} therefore takes this LLM-call
boundary as its uniform intervention point across backends. At each call,
the prompt is extended to request $K$ plausible answers with self-reported
preference weights, so a single sampled output is replaced by an explicit
candidate set. The backend then receives one final response and remains
agnostic to the watermark's presence. In effect, open-vocabulary
structured generation is recast as a finite, distribution-bearing choice
that can be keyed, audited, and replayed.

\paragraph{Self-reported candidate elicitation.}
Parsing the wrapper response yields a decision list
$\{d_t^1,\ldots,d_t^K\}$ with normalized weights, which induce the
explicit distribution $\probdist_t$ over the candidate support. This
reduces open-vocabulary generation to a discrete selection that a keyed
sampler can consume. 

\paragraph{Keyed distribution-preserving sampling.}
\memmark{} feeds $(\candset_t,\probdist_t,\ctx_t)$ to a
distribution-preserving binning sampler keyed by the secret $K$ and the
context-bound nonce $\nonceT$ from Section~\ref{sec:problem-channel}. The
sampler returns the selected candidate $\hat c_t$ together with the
number of embedded bits, while preserving the marginal:
\begin{equation}
\Pr[\hat c_t=c\mid \probdist_t,K]=\probdist_t(c),\quad \forall c\in\candset_t.
\end{equation}


\paragraph{Properties.}
The sampler provides the following guarantees; proofs are deferred to
Appendix~\ref{app:proofs}.

\begin{lemma}[Strict distribution preservation]
\label{lem:dist-preserve}
For any valid $\probdist_t$, the keyed pick $\hat c_t$ satisfies
$\Pr[\hat c_t=c^i\mid K,\ctx_t]_{\mathrm{marg}}=\probdist_t(c^i)$.
Because this marginal depends only on $\probdist_t$ and $K$, it holds
identically regardless of which backend produced $\candset_t$.
\end{lemma}

\begin{lemma}[Cascade composition]
\label{lem:cascade}
Keyed picks across multiple calls of a single event are independent
under distinct context-derived nonces, so expected embedding capacity
grows approximately additively in the number of valid decisions.
\end{lemma}


% ---------- 4.4 audit trace ----------
\subsection{Cryptographic Audit Trace}
\label{sec:method-audit}
To support forensic attribution beyond the original execution
environment, \memmark{} records a three-layer cryptographic trace that
serves the three regimes of Section~\ref{sec:problem-threat}.

\paragraph{Per-decision commitments.}
Each sampled decision produces a commitment
\[
\begin{aligned}
\commit_t = H\big(\,
& \ctx_t \,\|\, H(\candset_t) \,\|\, H(\probdist_t) \,\|\, \\
& \hat c_t \,\|\, \text{bits}_t \,\|\, \nonceT \,\big),
\end{aligned}
\]
with $H$ collision-resistant and inputs canonically serialized.

\paragraph{Per-session Merkle tree and signed anchor.}
The per-decision commitments in a session form a Merkle tree whose root
is sealed in a signed anchor
\[
\begin{aligned}
\header = (\,&\text{agent\_id},\ \text{user\_id},\ \text{session\_id},\ T,\\
            &\text{root}_T,\ \mathrm{sig}_K(\text{root}_T)),
\end{aligned}
\]
making the trace append-only and tamper-evident.

\paragraph{Per-leaf inclusion proofs.}
Each record carries its Merkle inclusion path at seal time, so a surviving
leaf can be checked directly against the anchored root without rebuilding
the log. This prevents structural attacks from collapsing verification
into an all-or-nothing rebuilt-root test.
