% ============================================================
% Related Work — main paper keeps only the two CLOSEST lines:
%   §2.1 Agent behavioral watermarking (mechanism-adjacent)
%   §2.2 Long-term memory systems + provenance (problem-adjacent)
% Everything else (RAG/corpus, LLM text, crypto primitives,
% memory-ops-as-actions extras) → Appendix \ref{app:related-ext}.
% ============================================================
\section{Related Work}
\label{sec:related}

\todo{One-sentence lede: \memmark{} sits at the intersection of two
research lines — behavioral watermarking for LLM agents (which gives
us the mechanism) and long-term memory systems with provenance
(which gives us the threat model). Extended coverage of RAG /
corpus-level watermarking, LLM text watermarking, and cryptographic
provenance primitives is deferred to
Appendix~\ref{app:related-ext}.}

% ============================================================
\subsection{Behavioral Watermarking for LLM Agents}
\label{sec:related-behavioral}
% ============================================================
\todo{Cite \agentmark~\cite{agentmark}, Agent
Guide~\cite{agentguide}, ActHook~\cite{acthook},
AGENTWM~\cite{agentwm}. All four embed at the visible-action layer
(planning, tool call, trajectory) and require the trajectory at
verify time. \memmark{} embeds at the latent state-evolution
decision layer, which survives when the trajectory is gone.}

\todo{Relation to \agentmark: we reuse its binning sampler as a
black-box component inside \S\ref{sec:method-sampler}; the
contribution is not the sampler but (a) where to intercept inside a
third-party memory SDK, (b) how to construct $(\candset, \probdist)$
from open-vocabulary structured-JSON LLM output, and (c) how to
compose audits across a cascade of intercepted calls per memory
event.}

\todo{Adjacent line — memory ops as actions
(\textsc{Memory-as-Action}~\cite{mem-as-action},
A-MAC~\cite{a-mac}). These optimize \emph{which} memory op to take;
\memmark{} embeds attribution \emph{into} the op. Orthogonal but
the closest neighbor on the "memory writes are first-class
decisions" axis.}

% ============================================================
\subsection{Long-Term Memory Systems and Provenance}
\label{sec:related-memory}
% ============================================================
\todo{Operating environment: \amem~\cite{amem},
\graphiti~\cite{graphiti}, MemOS~\cite{memos},
MemMachine~\cite{memmachine}. Two are the backends we wrap
(\amem{} agentic notes, \graphiti{} temporal graph); the other two
are MemMark-compatible via the §\ref{sec:method-formal} adapter.}

\todo{Provenance-aware designs:
\tiermem~\cite{tiermem} ships explicit provenance chains and
immutable source anchoring; MemOS \texttt{MemCube} carries
versioning / origin / lifecycle fields. Both are
\emph{field-stamping} provenance — they encode "who wrote this"
into metadata that the writer controls. \memmark{}'s contrast is
exactly this: we treat the snapshot-holder as the writer (adversarial
storage), so any field-based provenance is forgeable. \tiermem{} /
MemOS therefore appear in \S\ref{sec:rq3} as the
\texttt{signed-metadata-only} ablation point.}

\todo{Threat-side motivation: A-MemGuard~\cite{amemguard} and
MemoryGraft~\cite{memorygraft} catalog memory-poisoning attacks on
LLM agents — these motivate our \S\ref{sec:rq4} attack model
(\texttt{poisoning}, \texttt{manual\_edits}). The Mnemonic
Sovereignty survey~\cite{mnemonicsovereignty} covers governance /
access control / poisoning defense but explicitly does not cover
behavioral watermarking on memory-evolve decisions, supporting our
positioning.}

\todo{Direct KG-backend baseline: \kgmark~\cite{kgmark} is dynamic
knowledge-graph watermarking and is the closest single-point
comparison on KG-structured memory. \S\ref{sec:rq1}–\S\ref{sec:rq4}
compare \memmark{} vs \kgmark{} on the \graphiti{} backend.
\kgmark{} is not applicable to \amem{} (notes network, not strict
KG).}

% ---------- positioning table (stays in main paper) ----------
\input{tables/related_positioning}
